---
title: "Assignment 8"
author: "Dans Lismanis"
date: "2023-11-27"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(maptree)
library(janitor)
library(broom)
library(knitr)
library(tidyverse)
library(ggplot2)
library(zoo)
library(openxlsx)
library(xgboost)
set.seed(69)
```

## Question 1

### Importing data:
```{r}
oj <- read_csv('/Users/dans/Desktop/ECON/ass4/oj.csv',
               show_col_types = FALSE) %>% 
  mutate(logprice = log(price)) %>% 
  clean_names()

# create a wide dataset
oj_wide <- oj %>% 
  select(store, week, brand, logmove, logprice, feat) %>% 
  pivot_longer(logmove:feat) %>% 
  mutate(name = str_c(name, brand, sep = '_')) %>% 
  pivot_wider(id_cols = c(store, week), names_from = name, values_from = value)
 
# merge on the wide features a few times to create lags
oj_model <- oj %>% 
  left_join(oj_wide %>% 
              mutate(week = week + 1) %>% 
              rename_with(~(str_c(.x, '_lag_1')), .cols = c(logmove_minute.maid:feat_dominicks)),
            by = c('store', 'week')) %>% 
  left_join(oj_wide %>% 
              mutate(week = week + 2) %>% 
              rename_with(~(str_c(.x, '_lag_2')), .cols = c(logmove_minute.maid:feat_dominicks)),
            by = c('store', 'week')) %>% 
  left_join(oj_wide %>% 
              mutate(week = week + 3) %>% 
              rename_with(~(str_c(.x, '_lag_3')), .cols = c(logmove_minute.maid:feat_dominicks)),
            by = c('store', 'week')) %>% 
  mutate(brand = factor(brand),
         week = week - min(week),
         id = row_number()) %>%  # these identifiers are only necessary for the models where we hold out data, which aren't required
  select(-sstrvol, -cpwvol5) %>% # remove these features since they involve sales volumes, not sure how calculated
  na.omit()
```

#### For some reason a few columns have changed their names

```{r}
colnames(oj_model)[colnames(oj_model) == "logmove_tropicana.x"] <- "logmove_tropicana_lag_1"
colnames(oj_model)[colnames(oj_model) == "logprice_tropicana.x"] <- "logprice_tropicana_lag_1"
colnames(oj_model)[colnames(oj_model) == "feat_tropicana.x"] <- "feat_tropicana_lag_1"

colnames(oj_model)[colnames(oj_model) == "logmove_tropicana.y"] <- "logmove_tropicana_lag_2"
colnames(oj_model)[colnames(oj_model) == "logprice_tropicana.y"] <- "logprice_tropicana_lag_2"
colnames(oj_model)[colnames(oj_model) == "feat_tropicana.y"] <- "feat_tropicana_lag_2"

colnames(oj_model)[colnames(oj_model) == "logmove_tropicana"] <- "logmove_tropicana_lag_3"
colnames(oj_model)[colnames(oj_model) == "logprice_tropicana"] <- "logprice_tropicana_lag_3"
colnames(oj_model)[colnames(oj_model) == "feat_tropicana"] <- "feat_tropicana_lag_3"
```
### Running a tree :

```{r}
oj_tree <- oj_model %>% 
  mutate(q = exp(logmove)) %>% 
  group_by(store, week) %>% 
  mutate(weighted_mean = weighted.mean(price, q)) %>% 
  ungroup()
```

```{r}
reg_tree_data <- oj_tree %>% 
  select(weighted_mean, age60:cpdist5)
```

```{r}
fit<-rpart(as.formula(weighted_mean ~ .),data=reg_tree_data,method="anova",cp=0.007)
draw.tree((fit))
```
  
  draw.tree(fit)

### Assigning leafs to entries:
```{r}
ojleafs=oj_tree
ojleafs$leaf = fit$where
```

### Creating dataframes for each leaf:
```{r}
l1=ojleafs[ojleafs$leaf==2,]
l2=ojleafs[ojleafs$leaf==4,]
l3=ojleafs[ojleafs$leaf==5,]
```

### Creating a dataframe for each brand in each leaf

```{r}
doml1=l1[l1$brand=="dominicks",]
doml2=l2[l2$brand=='dominicks',]
doml3=l3[l3$brand=="dominicks",]

minl1=l1[l1$brand=="minute.maid",]
minl2=l2[l2$brand=='minute.maid',]
minl3=l3[l3$brand=="minute.maid",]

trol1=l1[l1$brand=="tropicana",]
trol2=l2[l2$brand=='tropicana',]
trol3=l3[l3$brand=="tropicana",]
```

### Getting residuals for each brand for each leaf:

#### Dominicks Price residuals leaf1

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(doml1)),size=(0.5*nrow(doml1)),replace = FALSE)
doml1Train=doml1[selcindex,]
doml1Test=doml1[-selcindex,]
formPricedoml1= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
PriceDoml1MTrain=model.matrix(formPricedoml1,data=doml1Train)
PriceDoml1MTest=model.matrix(formPricedoml1,data=doml1Test)
PriceDoml1DM=xgb.DMatrix(data=PriceDoml1MTrain, label=doml1Train$logprice)
crossvPdoml1=xgb.cv(data=PriceDoml1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPDoml1=xgboost(data=PriceDoml1DM, nrounds=102, print_every_n = 150)
predictedPDoml1=predict(XGBPDoml1,newdata=PriceDoml1MTest)
residualsPDoml1=doml1Test$logprice-predictedPDoml1
```

#### Dominicks Price residuals leaf2

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(doml2)),size=(0.5*nrow(doml2)),replace = FALSE)
doml2Train=doml2[selcindex,]
doml2Test=doml2[-selcindex,]
formPricedoml2= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
PriceDoml2MTrain=model.matrix(formPricedoml2,data=doml2Train)
PriceDoml2MTest=model.matrix(formPricedoml2,data=doml2Test)
PriceDoml2DM=xgb.DMatrix(data=PriceDoml2MTrain, label=doml2Train$logprice)
crossvPdoml2=xgb.cv(data=PriceDoml2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPDoml2=xgboost(data=PriceDoml2DM, nrounds=154, print_every_n = 150)
predictedPDoml2=predict(XGBPDoml2,newdata=PriceDoml2MTest)
residualsPDoml2=doml2Test$logprice-predictedPDoml2
```

#### Dominicks Price residuals leaf 3

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(doml3)),size=(0.5*nrow(doml3)),replace = FALSE)
doml3Train=doml3[selcindex,]
doml3Test=doml3[-selcindex,]
formPricedoml3= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
PriceDoml3MTrain=model.matrix(formPricedoml3,data=doml3Train)
PriceDoml3MTest=model.matrix(formPricedoml3,data=doml3Test)
PriceDoml3DM=xgb.DMatrix(data=PriceDoml3MTrain, label=doml3Train$logprice)
crossvPdoml3=xgb.cv(data=PriceDoml3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPDoml3=xgboost(data=PriceDoml3DM, nrounds=34, print_every_n = 150)
predictedPDoml3=predict(XGBPDoml3,newdata=PriceDoml3MTest)
residualsPDoml3=doml3Test$logprice-predictedPDoml3
```
#### Dominicks quantity residuals leaf 1

```{r}
set.seed(69)
formMovedoml1= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
MoveDoml1MTrain=model.matrix(formMovedoml1,data=doml1Train)
MoveDoml1MTest=model.matrix(formMovedoml1,data=doml1Test)
MoveDoml1DM=xgb.DMatrix(data=MoveDoml1MTrain, label=doml1Train$logmove)
crossvMovedoml1=xgb.cv(data=MoveDoml1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveDoml1=xgboost(data=MoveDoml1DM, nrounds=88, print_every_n = 150)
predictedMoveDoml1=predict(XGBMoveDoml1,newdata=MoveDoml1MTest)
residualsMoveDoml1=doml1Test$logmove-predictedMoveDoml1
```

#### Dominicks quantity residuals leaf 2

```{r}
set.seed(69)
formMovedoml2= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
MoveDoml2MTrain=model.matrix(formMovedoml2,data=doml2Train)
MoveDoml2MTest=model.matrix(formMovedoml2,data=doml2Test)
MoveDoml2DM=xgb.DMatrix(data=MoveDoml2MTrain, label=doml2Train$logmove)
crossvMovedoml2=xgb.cv(data=MoveDoml2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveDoml2=xgboost(data=MoveDoml2DM, nrounds=52, print_every_n = 150)
predictedMoveDoml2=predict(XGBMoveDoml2,newdata=MoveDoml2MTest)
residualsMoveDoml2=doml2Test$logmove-predictedMoveDoml2
```

#### Dominicks quantity residuals leaf 3

```{r}
set.seed(69)
formMovedoml3= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
MoveDoml3MTrain=model.matrix(formMovedoml3,data=doml3Train)
MoveDoml3MTest=model.matrix(formMovedoml3,data=doml3Test)
MoveDoml3DM=xgb.DMatrix(data=MoveDoml3MTrain, label=doml3Train$logmove)
crossvMovedoml3=xgb.cv(data=MoveDoml3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveDoml3=xgboost(data=MoveDoml3DM, nrounds=59, print_every_n = 150)
predictedMoveDoml3=predict(XGBMoveDoml3,newdata=MoveDoml3MTest)
residualsMoveDoml3=doml3Test$logmove-predictedMoveDoml3
```

#### Minute maid price leaf 1

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(minl1)),size=(0.5*nrow(minl1)),replace = FALSE)
minl1Train=minl1[selcindex,]
minl1Test=minl1[-selcindex,]
formPriceminl1= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Priceminl1MTrain=model.matrix(formPriceminl1,data=minl1Train)
Priceminl1MTest=model.matrix(formPriceminl1,data=minl1Test)
Priceminl1DM=xgb.DMatrix(data=Priceminl1MTrain, label=minl1Train$logprice)
crossvPminl1=xgb.cv(data=Priceminl1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPminl1=xgboost(data=Priceminl1DM, nrounds=97, print_every_n = 150)
predictedPminl1=predict(XGBPminl1,newdata=Priceminl1MTest)
residualsPminl1=minl1Test$logprice-predictedPminl1
```

#### Minute maid price leaf 2

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(minl2)),size=(0.5*nrow(minl2)),replace = FALSE)
minl2Train=minl2[selcindex,]
minl2Test=minl2[-selcindex,]
formPriceminl2= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Priceminl2MTrain=model.matrix(formPriceminl2,data=minl2Train)
Priceminl2MTest=model.matrix(formPriceminl2,data=minl2Test)
Priceminl2DM=xgb.DMatrix(data=Priceminl2MTrain, label=minl2Train$logprice)
crossvPminl2=xgb.cv(data=Priceminl2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPminl2=xgboost(data=Priceminl2DM, nrounds=150, print_every_n = 150)
predictedPminl2=predict(XGBPminl2,newdata=Priceminl2MTest)
residualsPminl2=minl2Test$logprice-predictedPminl2
```

#### Minute maid price leaf 3

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(minl3)),size=(0.5*nrow(minl3)),replace = FALSE)
minl3Train=minl3[selcindex,]
minl3Test=minl3[-selcindex,]
formPriceminl3= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Priceminl3MTrain=model.matrix(formPriceminl3,data=minl3Train)
Priceminl3MTest=model.matrix(formPriceminl3,data=minl3Test)
Priceminl3DM=xgb.DMatrix(data=Priceminl3MTrain, label=minl3Train$logprice)
crossvPminl3=xgb.cv(data=Priceminl3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPminl3=xgboost(data=Priceminl3DM, nrounds=81, print_every_n = 150)
predictedPminl3=predict(XGBPminl3,newdata=Priceminl3MTest)
residualsPminl3=minl3Test$logprice-predictedPminl3
```

#### Minute maid quantity residuals leaf 1

```{r}
set.seed(69)
formMoveminl1= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Moveminl1MTrain=model.matrix(formMoveminl1,data=minl1Train)
Moveminl1MTest=model.matrix(formMoveminl1,data=minl1Test)
Moveminl1DM=xgb.DMatrix(data=Moveminl1MTrain, label=minl1Train$logmove)
crossvMoveminl1=xgb.cv(data=Moveminl1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveminl1=xgboost(data=Moveminl1DM, nrounds=157, print_every_n = 150)
predictedMoveminl1=predict(XGBMoveminl1,newdata=Moveminl1MTest)
residualsMoveminl1=minl1Test$logmove-predictedMoveminl1
```

#### Minute maid quantity residuals leaf 2

```{r}
set.seed(69)
formMoveminl2= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Moveminl2MTrain=model.matrix(formMoveminl2,data=minl2Train)
Moveminl2MTest=model.matrix(formMoveminl2,data=minl2Test)
Moveminl2DM=xgb.DMatrix(data=Moveminl2MTrain, label=minl2Train$logmove)
crossvMoveminl2=xgb.cv(data=Moveminl2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveminl2=xgboost(data=Moveminl2DM, nrounds=57, print_every_n = 150)
predictedMoveminl2=predict(XGBMoveminl2,newdata=Moveminl2MTest)
residualsMoveminl2=minl2Test$logmove-predictedMoveminl2
```

#### Minute maid quantity residuals leaf 3

```{r}
set.seed(69)
formMoveminl3= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Moveminl3MTrain=model.matrix(formMoveminl3,data=minl3Train)
Moveminl3MTest=model.matrix(formMoveminl3,data=minl3Test)
Moveminl3DM=xgb.DMatrix(data=Moveminl3MTrain, label=minl3Train$logmove)
crossvMoveminl3=xgb.cv(data=Moveminl3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMoveminl3=xgboost(data=Moveminl3DM, nrounds=60, print_every_n = 150)
predictedMoveminl3=predict(XGBMoveminl3,newdata=Moveminl3MTest)
residualsMoveminl3=minl3Test$logmove-predictedMoveminl3
```
#### Tropicana Price residuals leaf1

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(trol1)),size=(0.5*nrow(trol1)),replace = FALSE)
trol1Train=trol1[selcindex,]
trol1Test=trol1[-selcindex,]
formPricetrol1= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Pricetrol1MTrain=model.matrix(formPricetrol1,data=trol1Train)
Pricetrol1MTest=model.matrix(formPricetrol1,data=trol1Test)
Pricetrol1DM=xgb.DMatrix(data=Pricetrol1MTrain, label=trol1Train$logprice)
crossvPtrol1=xgb.cv(data=Pricetrol1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPtrol1=xgboost(data=Pricetrol1DM, nrounds=474, print_every_n = 150)
predictedPtrol1=predict(XGBPtrol1,newdata=Pricetrol1MTest)
residualsPtrol1=trol1Test$logprice-predictedPtrol1
```

#### Tropicana Price residuals leaf 2

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(trol2)),size=(0.5*nrow(trol2)),replace = FALSE)
trol2Train=trol2[selcindex,]
trol2Test=trol2[-selcindex,]
formPricetrol2= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Pricetrol2MTrain=model.matrix(formPricetrol2,data=trol2Train)
Pricetrol2MTest=model.matrix(formPricetrol2,data=trol2Test)
Pricetrol2DM=xgb.DMatrix(data=Pricetrol2MTrain, label=trol2Train$logprice)
crossvPtrol2=xgb.cv(data=Pricetrol2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPtrol2=xgboost(data=Pricetrol2DM, nrounds=140, print_every_n = 150)
predictedPtrol2=predict(XGBPtrol2,newdata=Pricetrol2MTest)
residualsPtrol2=trol2Test$logprice-predictedPtrol2
```

#### Tropicana Price residuals leaf 3

```{r}
set.seed(69)
selcindex=sample(seq_len(nrow(trol3)),size=(0.5*nrow(trol3)),replace = FALSE)
trol3Train=trol3[selcindex,]
trol3Test=trol3[-selcindex,]
formPricetrol3= logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Pricetrol3MTrain=model.matrix(formPricetrol3,data=trol3Train)
Pricetrol3MTest=model.matrix(formPricetrol3,data=trol3Test)
Pricetrol3DM=xgb.DMatrix(data=Pricetrol3MTrain, label=trol3Train$logprice)
crossvPtrol3=xgb.cv(data=Pricetrol3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPtrol3=xgboost(data=Pricetrol3DM, nrounds=42, print_every_n = 150)
predictedPtrol3=predict(XGBPtrol3,newdata=Pricetrol3MTest)
residualsPtrol3=trol3Test$logprice-predictedPtrol3
```

#### Tropicana quantity residuals leaf 1

```{r}
set.seed(69)
formMovetrol1= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Movetrol1MTrain=model.matrix(formMovetrol1,data=trol1Train)
Movetrol1MTest=model.matrix(formMovetrol1,data=trol1Test)
Movetrol1DM=xgb.DMatrix(data=Movetrol1MTrain, label=trol1Train$logmove)
crossvMovetrol1=xgb.cv(data=Movetrol1DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMovetrol1=xgboost(data=Movetrol1DM, nrounds=121, print_every_n = 150)
predictedMovetrol1=predict(XGBMovetrol1,newdata=Movetrol1MTest)
residualsMovetrol1=trol1Test$logmove-predictedMovetrol1
```
#### Tropicana quantity residuals leaf 2

```{r}
set.seed(69)
formMovetrol2= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Movetrol2MTrain=model.matrix(formMovetrol2,data=trol2Train)
Movetrol2MTest=model.matrix(formMovetrol2,data=trol2Test)
Movetrol2DM=xgb.DMatrix(data=Movetrol2MTrain, label=trol2Train$logmove)
crossvMovetrol2=xgb.cv(data=Movetrol2DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMovetrol2=xgboost(data=Movetrol2DM, nrounds=138, print_every_n = 150)
predictedMovetrol2=predict(XGBMovetrol2,newdata=Movetrol2MTest)
residualsMovetrol2=trol2Test$logmove-predictedMovetrol2
```

#### Tropicana quantity residuals leaf 3

```{r}
set.seed(69)
formMovetrol3= logmove ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3
Movetrol3MTrain=model.matrix(formMovetrol3,data=trol3Train)
Movetrol3MTest=model.matrix(formMovetrol3,data=trol3Test)
Movetrol3DM=xgb.DMatrix(data=Movetrol3MTrain, label=trol3Train$logmove)
crossvMovetrol3=xgb.cv(data=Movetrol3DM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBMovetrol3=xgboost(data=Movetrol3DM, nrounds=43, print_every_n = 150)
predictedMovetrol3=predict(XGBMovetrol3,newdata=Movetrol3MTest)
residualsMovetrol3=trol3Test$logmove-predictedMovetrol3
```
##### Dominicks regressions

```{r}
regdoml1=lm(predictedMoveDoml1~residualsPDoml1+residualsPminl1+residualsPtrol1)
regdoml2=lm(predictedMoveDoml2~residualsPDoml2+residualsPminl2+residualsPtrol2)
regdoml3=lm(predictedMoveDoml3~residualsPDoml3+residualsPminl3+residualsPtrol3)
```

##### Minute maid regressions

```{r}
regminl1=lm(predictedMoveminl1~residualsPDoml1+residualsPminl1+residualsPtrol1)
regminl2=lm(predictedMoveminl2~residualsPDoml2+residualsPminl2+residualsPtrol2)
regminl3=lm(predictedMoveminl3~residualsPDoml3+residualsPminl3+residualsPtrol3)
```

##### Tropicana regressions

```{r}
regtrol1=lm(predictedMovetrol1~residualsPDoml1+residualsPminl1+residualsPtrol1)
regtrol2=lm(predictedMovetrol2~residualsPDoml2+residualsPminl2+residualsPtrol2)
regtrol3=lm(predictedMovetrol3~residualsPDoml3+residualsPminl3+residualsPtrol3)
```

#### Leaf 1 Matrix:

```{r}
cmat=matrix(coef(regdoml1),nrow=1)
cmat=rbind(cmat,coef(regminl1))
cmat=rbind(cmat,coef(regtrol1))
cmat=cmat[,-1]
rownames(cmat)=c("LnPriceDom","LnPriceMin","LnPriceTro")
cmat
```
#### Leaf 2 Matrix:

```{r}
cmat=matrix(coef(regdoml2),nrow=1)
cmat=rbind(cmat,coef(regminl2))
cmat=rbind(cmat,coef(regtrol2))
cmat=cmat[,-1]
rownames(cmat)=c("LnPriceDom","LnPriceMin","LnPriceTro")
cmat
```

#### Leaf 3 Matrix:

```{r}
cmat=matrix(coef(regdoml3),nrow=1)
cmat=rbind(cmat,coef(regminl3))
cmat=rbind(cmat,coef(regtrol3))
cmat=cmat[,-1]
rownames(cmat)=c("LnPriceDom","LnPriceMin","LnPriceTro")
cmat
```
### c

We compare XGBOOST with OLS for leaf 1 dominicks regression.
```{r}
#Using OLS to get price residuals
onec=lm(logprice ~ logmove_tropicana_lag_1+logprice_tropicana_lag_1+feat_tropicana_lag_1+logmove_minute.maid_lag_1+logprice_minute.maid_lag_1+feat_minute.maid_lag_1+logmove_dominicks_lag_1+logprice_dominicks_lag_1+feat_dominicks_lag_1+logmove_tropicana_lag_2+logprice_tropicana_lag_2+feat_tropicana_lag_2+logmove_minute.maid_lag_2+logprice_minute.maid_lag_2+feat_minute.maid_lag_2+logmove_dominicks_lag_2+logprice_dominicks_lag_2+feat_dominicks_lag_2+logmove_tropicana_lag_3+logprice_tropicana_lag_3+feat_tropicana_lag_3+logmove_minute.maid_lag_3+logprice_minute.maid_lag_3+feat_minute.maid_lag_3+logmove_dominicks_lag_3+logprice_dominicks_lag_3+feat_dominicks_lag_3,data=doml1Train)
predonec=predict(onec,newdata=doml1Test)
resdonec=doml1Test$logprice-predonec

```

```{r}
#Plotting residuals
plot(predict(onec), resid(onec), main = "Scatter Plot of OLS Residuals", 
     xlab = "Independent Variable", ylab = "Residuals")

plot(predictedPDoml1,residualsPDoml1 , main = "Scatter Plot of XGBOOST Residuals", 
     xlab = "Independent Variable", ylab = "Residuals")

```
FOR XGBOOST rediduals are much closer to 0 so it preforms better.

## Question 2

### a
They make 10 per user => price of high is 15
N=1mil, a=0.3, b=0, $$m_l$$=0, $$p_h$$=10
```{r}
Revenue=((1000000*0.3*0.5))*10*(12-2)
Revenue
```

### b

```{r}
Costs=(1000000*0.3)*0.5*12
Costs
```

### c

```{r}
ProfitNoFix=Revenue-Costs
ProfitNoFix
ProfitFix=Revenue-Costs-1000000
ProfitFix
```

### d

Since new revenue is less than the previous one they should not implement this change.
```{r}
NewRevenue=1000000*0.28*0.45*10*(12-1)
NewCosts=1000000*0.28*0.5*12
NewGrossMarg=NewRevenue-NewCosts
NewGrossMarg
```
## Question 3

#### Reading data

```{r}
ordata=read.xlsx("/Users/dans/Desktop/ECON/online_retail.xlsx")
ordata$InvoiceDate=convertToDate(ordata$InvoiceDate)
```

#### Manipulating data

```{r}
orders=ordata
orders$Revenue=orders$UnitPrice*orders$Quantity
orders <- orders %>%
  group_by(InvoiceNo,InvoiceDate,CustomerID,Country) %>%
  mutate(OrderTotal = sum(Revenue))
cleared <- orders %>%
  distinct(InvoiceNo,InvoiceDate,CustomerID,Country,OrderTotal)
#Removing rows with no customerID
cleared=cleared[!is.na(cleared$CustomerID),]
```

### a Summary statistics:

#### Unique customers:
```{r}
custID=cleared$CustomerID
NumUniqueCust=length((unique(custID)))
NumUniqueCust
```

#### Unique countries: 
```{r}
contrID=cleared$Country
NumUniqueCountr=length((unique(contrID)))
NumUniqueCountr
```

#### Total Revenue for each country:
```{r}
countrytot=cleared%>% 
  group_by(Country) %>%
  summarize(Total_Revenue=sum(OrderTotal))
countrytot
```

#### Total revenue per customer:
```{r}
custtot=cleared%>% 
  group_by(CustomerID) %>%
  summarize(Total_Revenue=sum(OrderTotal))
custtot
```

#### Average per country:

```{r}
countryavg=cleared%>% 
  group_by(Country) %>%
  summarize(Average_Revenue=mean(OrderTotal))
countryavg
```

#### Average per customer:

```{r}
custavg=cleared%>% 
  group_by(CustomerID) %>%
  summarize(Average_Revenue=mean(OrderTotal))
custavg
```

#### Average per customer in each country:

```{r}
contrcustavg=cleared%>% 
  group_by(Country,CustomerID) %>%
  summarize(Average_Revenue=mean(OrderTotal))
contrcustavg
```

### (b) 

#### Ploting average per customer:

```{r}
ggplot(custavg, aes(x=Average_Revenue)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.1,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot
```
#### Ploting total per customer:

```{r}
ggplot(custtot, aes(x=Total_Revenue)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.1,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot
```

### c

```{r}
#Sorting data in descending order
sortot=custtot[order(-custtot$Total_Revenue),]
sortot$cumilative_sum=cumsum(sortot$Total_Revenue)
eightyperc=which(sortot$cumilative_sum>=0.8*sum(sortot$Total_Revenue))[1]
eightyperc
```



### d

```{r}
df=cleared
# Convert 'invoice_date' to a date type if it's not already
df$invoice_date <- as.Date(df$InvoiceDate)

df <- df %>% filter(!grepl("^C", InvoiceNo, ignore.case = TRUE))
# Sort the data frame by 'customerID' and 'invoice_date'
df <- df %>% arrange(CustomerID, InvoiceDate)

# Create a new column 'next_invoice_date' with the next invoice date for each customer
df <- df %>% group_by(CustomerID) %>% mutate(next_invoice_date = lead(InvoiceDate))

# If you want to calculate the time difference between consecutive invoices, you can also create a 'time_difference' column
df <- df %>% mutate(time_difference = next_invoice_date - invoice_date)

df <- df %>% mutate(Month = format(invoice_date, "%Y-%m"))

# Count unique customers in each month
unique_customers_per_month <- df %>% group_by(Month) %>% summarise(unique_customers = n_distinct(CustomerID))

# Identify churners for each month (assuming churners are customers with a time_difference more than 3 months)
churners_per_month <- df %>%
  group_by(Month) %>%
  summarise(churners = sum(is.na(time_difference) | abs(time_difference) > 90))

UniqueANDChurners <- merge(unique_customers_per_month, churners_per_month, by = "Month", all.x = TRUE)

UniqueANDChurners$churn_rate=UniqueANDChurners$churners/UniqueANDChurners$unique_customers
UniqueANDChurners <- UniqueANDChurners[-nrow(UniqueANDChurners), ]
UniqueANDChurners

df$Churner <- ifelse(is.na(df$time_difference) | abs(df$time_difference) > 90, 1, 0)

df$BoughtAgain <- ifelse(abs(df$time_difference) > 90, 1, 0)

returned=sum(df$BoughtAgain == 1, na.rm = TRUE)
numchurn=sum(UniqueANDChurners$churners)
returned/numchurn

chrate=mean(UniqueANDChurners$churn_rate)
chrate
```
### e

```{r}
10/chrate
```

### f
```{r}
returns=cleared[cleared$OrderTotal<0,]
percreturned=sum(cleared$OrderTotal<0)/sum(cleared$OrderTotal>0)
percreturned
percrevreturned=abs(sum(cleared$OrderTotal[cleared$OrderTotal<0])/sum(cleared$OrderTotal[cleared$OrderTotal>0]))
percrevreturned
```

## Question 4

### a

```{r}
cleared <- cleared %>%
  mutate(day = format(InvoiceDate, "%d"),
         month = format(InvoiceDate, "%m"),
         year = format(InvoiceDate, "%Y"))

mean_revenue_per_day <- cleared %>%
  group_by(day) %>%
  summarise(mean_revenue = mean(OrderTotal, na.rm = TRUE))

mean_revenue_per_month <- cleared %>%
  group_by(month) %>%
  summarise(mean_revenue = mean(OrderTotal, na.rm = TRUE))

mean_revenue_per_year <- cleared %>%
  group_by(year) %>%
  summarise(mean_revenue = mean(OrderTotal, na.rm = TRUE))

mean_revenue_per_day
mean_revenue_per_month
mean_revenue_per_year
```

### b

```{r}
#For each day
 ggplot(mean_revenue_per_day, aes(x = day, y = mean_revenue)) +
  geom_point() +
  labs(title = "Mean Revenue for each day of month",
       x = "Day",
       y = "Revenue")

ggplot(mean_revenue_per_month, aes(x = month, y = mean_revenue)) +
  geom_point() +
  labs(title = "Mean Revenue for each month",
       x = "Month",
       y = "Revenue")

ggplot(mean_revenue_per_year, aes(x = year, y = mean_revenue)) +
  geom_point() +
  labs(title = "Mean Revenue for each Year",
       x = "Year",
       y = "Revenue")

```

### c

I would disagree because it might be that holidays themsleves cause spikes in sales.

## Question 5

### a
```{r}
FirstOrder= cleared %>%
  group_by(CustomerID) %>%
  summarize(first_order_date = min(InvoiceDate))
cleared=left_join(cleared,FirstOrder,by="CustomerID")
cleared$first_order_date=format(cleared$first_order_date, "%Y-%m")

```

### b

```{r}
FIVEa=cleared%>% 
  group_by(first_order_date) %>%
  summarize(MeanOrderTotal=mean(OrderTotal),
            OrderCount=sum(OrderTotal>0),
            ReturnCount=sum(OrderTotal<0),
            TotalOrderValue=sum(OrderTotal[OrderTotal>0]),
            TotalreturnValue=sum(OrderTotal[OrderTotal<0]),
            RealOrderValue=sum(OrderTotal)
            
            
            )
FIVEa
```
### c

We can see that there was a peak of Revenue at the start date and during the 2 month trial, after it dropped. If we introduced low level of subscription some of those users might be retained. 