---
title: "Assignment7"
author: "Dans Lismanis"
date: "2023-11-13"
output: html_document
---
```{r setup}
library(dplyr)
library(ggplot2)
library(xgboost)
set.seed(69)
```
## Question 1

### a

This firm runs the risk of losing the customers with high quality preference.

### b
1) This firm could produce two types of product, one for the high preference and one for the low preference customers.

2) This firm can do sales to cater to low quality preference which is usually due to the lower price

### c

Denote: PS0 - firms surplus now
        PS1- firms surplus with price descrimination
        WP- willingness to pay to identify customers:
        
WP<PS1-PS0

## Question 2

### a

#### Creating dataframes:

```{r}
#Creating dataframes for Dominicks,Minute Maid and Tropicana
df=read.csv("/Users/dans/Desktop/ECON/assignment1econ/oj.csv")
dfdom=df[df$brand=='dominicks',]
dfmin=df[df$brand=="minute.maid",]
dftro=df[df$brand=="tropicana",]

#Adding cross prices
dfdom$price=log(dfdom$price)
dfdom$LnPMin=log(dfmin$price)
dfdom$LnPTro=log(dftro$price)

dfmin$price=log(dfmin$price)
dfmin$LnPDom=dfdom$price
dfmin$LnPTro=log(dftro$price)

dftro$price=log(dftro$price)
dftro$LnPDom=dfdom$price
dftro$LnPMin=dfmin$price

# Creating Lags
df1dom=dfdom
df1dom$week=df1dom$week+1
dfdomWL=merge(dfdom,df1dom, by=c("brand","store","week"))
names(dfdomWL)=c("brand","store","week","logmove","feat","LnPrice","AGE60","EDUC","ETHNIC","INCOME","HHLARGE","WORKWOM","HVAL150","SSTRDIST","SSTRVOL","CPDIST5","CPWVOL5","LnPMin","LnPTro","Laglogmove","Lagfeat","LagLnPrice","LagAGE60","LagEDUC","LagETHNIC","LagINCOME","LagHHLARGE","LagWORKWOM","LagHVAL150","LagSSTRDIST","LagSSTRVOL","LagCPDIST5","LagCPWVOL5","LagLnPMin","LagLnPTro")

df1min=dfmin
df1min$week=df1min$week+1
dfminWL=merge(dfmin,df1min, by=c("brand","store","week"))
names(dfminWL)=c("brand","store","week","logmove","feat","LnPrice","AGE60","EDUC","ETHNIC","INCOME","HHLARGE","WORKWOM","HVAL150","SSTRDIST","SSTRVOL","CPDIST5","CPWVOL5","LnPDom","LnPTro","Laglogmove","Lagfeat","LagLnPrice","LagAGE60","LagEDUC","LagETHNIC","LagINCOME","LagHHLARGE","LagWORKWOM","LagHVAL150","LagSSTRDIST","LagSSTRVOL","LagCPDIST5","LagCPWVOL5","LagLnPDom","LagLnPTro")

df1tro=dftro
df1tro$week=df1tro$week+1
dftroWL=merge(dftro,df1tro, by=c("brand","store","week"))
names(dftroWL)=c("brand","store","week","logmove","feat","LnPrice","AGE60","EDUC","ETHNIC","INCOME","HHLARGE","WORKWOM","HVAL150","SSTRDIST","SSTRVOL","CPDIST5","CPWVOL5","LnPDom","LnPMin","Laglogmove","Lagfeat","LagLnPrice","LagAGE60","LagEDUC","LagETHNIC","LagINCOME","LagHHLARGE","LagWORKWOM","LagHVAL150","LagSSTRDIST","LagSSTRVOL","LagCPDIST5","LagCPWVOL5","LagLnPDom","LagLnPMin")
```

```{r}
#creating interactions:
dfdomWL$pAGE=dfdomWL$LagLnPrice*dfdomWL$LagAGE60
dfdomWL$pEDU=dfdomWL$LagLnPrice*dfdomWL$LagEDUC
dfdomWL$pETHNIC=dfdomWL$LagLnPrice*dfdomWL$LagETHNIC
dfdomWL$pWORKWOM=dfdomWL$LagLnPrice*dfdomWL$LagWORKWOM
dfdomWL$pHHLARGE=dfdomWL$LagLnPrice*dfdomWL$LagHHLARGE

dfminWL$pAGE=dfminWL$LagLnPrice*dfminWL$LagAGE60
dfminWL$pEDU=dfminWL$LagLnPrice*dfminWL$LagEDUC
dfminWL$pETHNIC=dfminWL$LagLnPrice*dfminWL$LagETHNIC
dfminWL$pWORKWOM=dfminWL$LagLnPrice*dfminWL$LagWORKWOM
dfminWL$pHHLARGE=dfminWL$LagLnPrice*dfminWL$LagHHLARGE

dftroWL$pAGE=dftroWL$LagLnPrice*dftroWL$LagAGE60
dftroWL$pEDU=dftroWL$LagLnPrice*dftroWL$LagEDUC
dftroWL$pETHNIC=dftroWL$LagLnPrice*dftroWL$LagETHNIC
dftroWL$pWORKWOM=dftroWL$LagLnPrice*dftroWL$LagWORKWOM
dftroWL$pHHLARGE=dftroWL$LagLnPrice*dftroWL$LagHHLARGE
```

#### For Dominicks:

Predicting Price And Quantity based on lagged variables:
```{r}
set.seed(69)
#For dominicks Price:
selcindex=sample(seq_len(nrow(dfdomWL)),size=(0.5*nrow(dfdomWL)),replace = FALSE)
dfdomWLTrain=dfdomWL[selcindex,]
dfdomWLTest=dfdomWL[-selcindex,]
formPricedom= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPMin + LagLnPTro + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
PriceDomMTrain=model.matrix(formPricedom,data=dfdomWLTrain)
PriceDomMTest=model.matrix(formPricedom,data=dfdomWLTest)
PriceDomDM=xgb.DMatrix(data=PriceDomMTrain, label=dfdomWLTrain$LnPrice)
crossvPdom=xgb.cv(data=PriceDomDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPDom=xgboost(data=PriceDomDM, nrounds=189, print_every_n = 150)
predictedPDom=predict(XGBPDom,newdata=PriceDomMTest)
residualsPDom=dfdomWLTest$LnPrice-predictedPDom
```

```{r}
set.seed(69)
#Finding residuals of Dom on LogMove:
formmovedom= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPMin + LagLnPTro + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
MoveDomMTrain=model.matrix(formmovedom,data=dfdomWLTrain)
MoveDomMTest=model.matrix(formmovedom,data=dfdomWLTest)
MoveDomDM=xgb.DMatrix(data=MoveDomMTrain, label=dfdomWLTrain$logmove)
crossvPdom=xgb.cv(data=MoveDomDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBQDom=xgboost(data=MoveDomDM, nrounds=246, print_every_n = 150)
predictedQDom=predict(XGBQDom,newdata=MoveDomMTest)
residualsQDom=dfdomWLTest$logmove-predictedQDom
```

OLS of resdPrice on resdQuan
```{r}
DomRes=data.frame(
  ResdLnMove=residualsQDom,
  ResdLnPrice=residualsPDom
)
DomResQuanOnPrice=lm(residualsQDom~residualsPDom,data=DomRes)
summary(DomResQuanOnPrice)
```

#### For Minute Maid:

Predicting Price And Quantity based on lagged variables:
```{r}
set.seed(69)
#For Minute Maid Price:
selcindex=sample(seq_len(nrow(dfminWL)),size=(0.5*nrow(dfminWL)),replace = FALSE)
dfminWLTrain=dfminWL[selcindex,]
dfminWLTest=dfminWL[-selcindex,]
formPricemin= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPDom + LagLnPTro + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
PriceminMTrain=model.matrix(formPricemin,data=dfminWLTrain)
PriceminMTest=model.matrix(formPricemin,data=dfminWLTest)
PriceminDM=xgb.DMatrix(data=PriceminMTrain, label=dfminWLTrain$LnPrice)
crossvPmin=xgb.cv(data=PriceminDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPmin=xgboost(data=PriceminDM, nrounds=255, print_every_n = 150)
predictedPmin=predict(XGBPmin,newdata=PriceminMTest)
residualsPmin=dfminWLTest$LnPrice-predictedPmin
```

```{r}
set.seed(69)
#Finding residuals of Minute Maid on LogMove:
formmovemin= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPDom + LagLnPTro + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
MoveminMTrain=model.matrix(formmovemin,data=dfminWLTrain)
MoveminMTest=model.matrix(formmovemin,data=dfminWLTest)
MoveminDM=xgb.DMatrix(data=MoveminMTrain, label=dfminWLTrain$logmove)
crossvPmin=xgb.cv(data=MoveminDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBQmin=xgboost(data=MoveminDM, nrounds=241, print_every_n = 150)
predictedQmin=predict(XGBQmin,newdata=MoveminMTest)
residualsQmin=dfminWLTest$logmove-predictedQmin
```

```{r}
minRes=data.frame(
  ResdLnMove=residualsQmin,
  ResdLnPrice=residualsPmin
)
minResQuanOnPrice=lm(residualsQmin~residualsPmin,data=minRes)
summary(minResQuanOnPrice)
```

#### For Tropicana:

Predicting Price And Quantity based on lagged variables:
```{r}
set.seed(69)
#For Tropicana Price:
selcindex=sample(seq_len(nrow(dftroWL)),size=(0.5*nrow(dftroWL)),replace = FALSE)
dftroWLTrain=dftroWL[selcindex,]
dftroWLTest=dftroWL[-selcindex,]
formPricetro= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPDom + LagLnPMin + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
PricetroMTrain=model.matrix(formPricetro,data=dftroWLTrain)
PricetroMTest=model.matrix(formPricetro,data=dftroWLTest)
PricetroDM=xgb.DMatrix(data=PricetroMTrain, label=dftroWLTrain$LnPrice)
crossvPtro=xgb.cv(data=PricetroDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBPtro=xgboost(data=PricetroDM, nrounds=242, print_every_n = 150)
predictedPtro=predict(XGBPtro,newdata=PricetroMTest)
residualsPtro=dftroWLTest$LnPrice-predictedPtro
```

```{r}
set.seed(69)
#Finding residuals of Tropicana on LogMove:
formmovetro= LnPrice ~ week + store + Laglogmove+ Lagfeat +LagLnPrice + LagAGE60 + LagEDUC + LagETHNIC + LagINCOME + LagHHLARGE + LagWORKWOM + LagHVAL150 + LagSSTRDIST + LagSSTRVOL + LagCPDIST5 + LagCPWVOL5 + LagLnPDom + LagLnPMin + pAGE + pEDU + pETHNIC + pWORKWOM + pHHLARGE
MovetroMTrain=model.matrix(formmovetro,data=dftroWLTrain)
MovetroMTest=model.matrix(formmovetro,data=dftroWLTest)
MovetroDM=xgb.DMatrix(data=MovetroMTrain, label=dftroWLTrain$logmove)
crossvPtro=xgb.cv(data=MovetroDM, nrounds = 1000, nfold=2, early_stopping_rounds = 150, print_every_n = 150)
XGBQtro=xgboost(data=MovetroDM, nrounds=112, print_every_n = 150)
predictedQtro=predict(XGBQtro,newdata=MovetroMTest)
residualsQtro=dftroWLTest$logmove-predictedQtro
```

```{r}
troRes=data.frame(
  ResdLnMove=residualsQtro,
  ResdLnPrice=residualsPtro
)
troResQuanOnPrice=lm(residualsQtro~residualsPtro,data=troRes)
summary(troResQuanOnPrice)
```

### b
```{r}
#creating a dataframe with all residuals:
dfresiduals=data.frame(
  LnMoveDom=residualsQDom,
  LnPriceDom=residualsPDom,
  LnMoveMin=residualsQmin,
  LnPriceMin=residualsPmin,
  LnMoveTro=residualsQtro,
  LnPriceTro=residualsPtro
  
)
```

```{r}
#OLS regresion for Dom:
regcrossdom=lm(LnMoveDom~LnPriceDom+LnPriceMin+LnPriceTro,data=dfresiduals)
summary(regcrossdom)
```

```{r}
#OLS regresion for Min:
regcrossmin=lm(LnMoveMin~LnPriceDom+LnPriceMin+LnPriceTro,data=dfresiduals)
summary(regcrossmin)
```

```{r}
#OLS regresion for Dom:
regcrosstro=lm(LnMoveTro~LnPriceDom+LnPriceMin+LnPriceTro,data=dfresiduals)
summary(regcrosstro)
```

```{r}
#Creating Matrix:
cmat=matrix(coef(regcrossdom),nrow=1)
cmat=rbind(cmat,coef(regcrossmin))
cmat=rbind(cmat,coef(regcrosstro))
cmat=cmat[,-1]
rownames(cmat)=c("LnPriceDom","LnPriceMin","LnPriceTro")
cmat
```